\documentclass[12pt]{letter}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}

\address{
  Abdulkadir Elmas, PhD\\
  Department of Systems Biology\\
  Columbia University Irving Medical Center\\
  1130 St.\ Nicholas Avenue\\
  New York, NY 10032, USA\\
  ae2321@columbia.edu
}

\date{\today}

\begin{document}

\begin{letter}{
  The Editors\\
  \textit{GigaScience}\\
  Oxford University Press
}

\opening{Dear Editors,}

I am writing to submit for your consideration the manuscript entitled \textbf{``BEACON-IO: Bayesian estimation of expression-driven cancer dependencies at the tumor-immune interface reveals rational immunotherapy combination targets''} for publication in \textit{GigaScience}.

This manuscript presents BEACON-IO, a computational framework that extends the recently published BEACON method (Elmas, \textit{GigaScience} 2026, giag011) from pan-cancer dependency discovery to precision immuno-oncology. The work addresses a pressing unmet need: identifying functional, druggable targets that explain why tumors resist immune checkpoint blockade (ICB) and what to combine with ICB to overcome resistance.

\textbf{The key contributions of this work are:}

\begin{enumerate}
  \item \textbf{Methodological innovation.} BEACON-IO is the first framework to systematically identify cancer dependencies that are modulated by the tumor immune microenvironment. Using Bayesian MCMC inference with the No-U-Turn Sampler, we quantify expression-dependency correlations with full posterior distributions and 95\% highest-density intervals---providing principled uncertainty quantification absent from prior frequentist approaches. Convergence is confirmed by $\hat{R} = 1.00$ and ESS $> 1{,}000$ across all reported results.

  \item \textbf{New biology.} Across 22 cancer lineages in DepMap 24Q4 (1,178 cell lines), we identify 166 EDD genes and 77 immune-context-specific dependencies. The top immune-hot specific targets---\textit{IRF4}, \textit{IKZF1}, \textit{MYB}, \textit{IKZF3}---converge on lymphoid identity programmes and the molecular targets of IMiDs (immunomodulatory drugs). This directly predicts that lenalidomide/thalidomide analogs may synergize with ICB in immune-infiltrated solid tumors, a hypothesis now testable in ongoing clinical trials.

  \item \textbf{Multi-scale clinical validation.} Integration of PRISM drug sensitivity (857 compounds, 23,372 significant gene-drug pairs), TCGA survival analysis (10 cancer types, 350 significant gene-cancer associations), and ICB response prediction (BEACON-IO AUC = 0.583 vs.\ 0.500 for GEP, CYT, and IMPRES in Hugo 2016 melanoma) provides converging evidence for top targets.

  \item \textbf{Full reproducibility.} The entire analysis pipeline---from data download through figure generation---is implemented in Python, orchestrated by Snakemake, and deposited at \url{https://github.com/aelmas/beacon-io} (MIT license). All data are obtained exclusively from public sources (DepMap, TCGA, PRISM, GEO). This is directly aligned with \textit{GigaScience}'s commitment to open, reproducible science.
\end{enumerate}

This work is directly relevant to \textit{GigaScience}'s scope as a data-intensive computational biology study: it processes $> 4$ GB of public data, produces a fully documented analysis pipeline, and deposits all processed results and figures in the GitHub repository. The methods are novel, the results are reproducible, and the biological findings have immediate translational implications.

I confirm that this manuscript has not been published elsewhere and is not under consideration by another journal. The work presents original research; the BEACON-IO framework is conceptually and technically distinct from the previously published BEACON paper, which focused on pan-cancer target discovery without immune context.

I declare no competing interests. This work received no specific funding.

I would be happy to suggest reviewers with expertise in computational oncology, cancer genomics, and immuno-oncology upon request.

Thank you for considering this manuscript.

\closing{Sincerely,}

\vspace{2cm}
Abdulkadir Elmas, PhD\\
Department of Systems Biology\\
Columbia University Irving Medical Center\\
ae2321@columbia.edu\\
\url{https://github.com/aelmas}

\end{letter}
\end{document}
